---
title: "Linear Regression for Photosynthesis Data"
output:
  md_document:
    variant: markdown_github
---

# Linear Regression for Photosynthesis Data

The file `photo.csv` contains data on the net maximal rate of photosynthesis of 198 plant species together with 4 other attributes related to their leaves. These are real, field data.

The variables are:

- Code: a code denoting the species
- LeafType: type of leaf (sunleaf or whole plant) 
- Amax: the net maximal rate of photosynthesis (A=assimilation) 
- LDMC: concentration of dry matter in the leaf 
- SLA: specific surface area of the leaf
- LNC: concentration of organic nitrogen in the leaf 

## Objective

Perform a simple linear regression to predict the net rate of photosynthesis (Amax) given the specific leaf area (SLA). Of course, many other models are possible... 

```{r}
# Read in the  data
photo <- read.csv("./photo.csv",header=T)
# display variable names and first lines
names(photo)
head(photo)
```

We perform some basic EDA

- a scatter plot of the explanatory and the response varible, colored by leaf type
- histograms of the variables of interest


```{r}
library(ggplot2)
ggplot(photo, aes(x=SLA, y=Amax, shape =LeafType, color=LeafType)) +
       geom_point()
ggplot(photo, aes(x=SLA)) +
    geom_histogram(binwidth=5, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
    ggtitle("SLA")

ggplot(photo, aes(x=Amax)) +
    geom_histogram(binwidth=40, fill="red", color="#e9ecef", alpha=0.9) +
    ggtitle("Amax")

```

There is a clearly a correlation between the two, with whole plants having higher values, in general..

Summary statistics:

```{r}
# Summary overview of `photo`
summary(photo) 
```

There are some missing values here, but they will be automatically treated by the linear regression.

Let us perform an initial linear regression of Amax as a function of SLA, and inspect the results.

```{r}
lr1 <- lm(Amax ~ SLA, data=photo, na.action=na.omit)
#par(mfrow=c(2,2))
plot(lr1)
summary(lr1)
```
## Analysis of model `lr1`

As expected from the scatter plot, the results are not very satisfactory. In spite of highly significant coefficients, the R-squared value is low and the diagnostic plots show:

1. A lack of homoscedasticity.
2. A clear deviation of the residulas from a Gaussian distributioN

This suggests that we try a nonlinear change of variables to better capture the biological phenomenon. 

One possibility is to regress the logarithm of Amax on the logarithm of SLA. This intuition come from an allometric law that is known to describe many biological processes.

```{r}
# Fit an allometric relation
#.    ln(y) = a + b*ln(x)
lr2<-lm(log(Amax)~log(SLA),data=photo,na.action=na.omit)
plot(lr2)
summary(lr2)
```

## Conclusions

- The allometric relation has improved the model.
- The value of R-squared has increased from 0.6135 to 0.7266
- The residuals are more homogeneously distributed.

